---
title: "Machine_Learning_Assign1"
author: "Martin Guo"
date:  18 Sep 2015
output: html_document
---

# Introduction #

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect 
a large amount of data about personal activity relatively inexpensively. These type of devices 
are part of the quantified self movement ?C a group of enthusiasts who take measurements 
about themselves regularly to improve their health, to find patterns in their behavior, or 
because they are tech geeks. One thing that people regularly do is quantify how much of a 
particular activity they do, but they rarely quantify how well they do it. In this project, 
our goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 
6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different 
ways. More information is available from the website here: 

http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

# Explore Data #

## Load Data##
```{r Load Explore , echo=TRUE}

train_data <- read.csv("./data/pml-training.csv", na.strings=c("NA","", NULL))
test_data <- read.csv("./data/pml-testing.csv", na.strings=c("NA","", NULL))

#See dimension of data
dim(train_data)
dim(test_data)

```

## Reduce Data##
There are many columns contains too many NA value, we remove these features so to 
decrease compute complexity. We set 10% as a threshold, many column whose NA rate
above it will be removed from data set.

```{r Reduce Explore , echo=TRUE}

# Remove appearently unrelated features 
train_data <- subset(train_data, select = -c(X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp))

# Remove columns which include too many NA vaclue
feature_na_num <- apply(train_data, 2, function(x) { sum(is.na(x))} )
many_na_column <- feature_na_num[which(feature_na_num > (dim(train_data)[1] * 0.10))]

train_data <- train_data[, -which(names(train_data) %in% names(many_na_column) )]

#Reduced data
dim(train_data)

#Remove highly correlated column

library(caret)

# filter factor columns 
cor_data <- subset(train_data, select = -c(new_window, classe))

cor_matrix <- cor(cor_data)

# get high correlation columns
highly_cor <- findCorrelation(cor_matrix, cutoff=0.7)

train_data <- subset(train_data, select = ! names(train_data) %in% names(cor_data)[highly_cor])

dim(train_data)
```

After data processing, we got 19622 samples , 33 features. By features reducing,
we finally got 33 features instead of 160 at the begining. The reducing will help
us cut machining learning time consuming greatly.



## Splite Data##
We will splite all samples into 2 groups, training (70%) and testing (30%) data. 
So that we can get Out-of-Sample error.
 
```{r Splite Data , echo=TRUE}

#split data into training data nd testing data
inTrain <- createDataPartition(y=train_data$classe, p=0.7, list=FALSE)

training <- train_data[inTrain,]
testing <- train_data[-inTrain,]

dim(training)
dim(testing)
```
Training data include : 13737 samples.
Testing data include : 5885 samples.

# Machine Learning #
We use Random Forest model for machine learning. Here we set number of tree to 100
in case of R generating too many tree to cost a very long time. We can tune this 
parameter according predicting accuracy and time consuming.
```{r Machine Learning , echo=TRUE}
library(e1071)
library(mboost)
set.seed(1024)

mod_rf <- train(classe ~ ., method="rf", data=training, ntree=100,
                  trControl = trainControl(method="cv"), number=3)

```

# Cross Validation Test #
```{r Cross Validation , echo=TRUE}
pred_rf <- predict(mod_rf, testing)

cm <- confusionMatrix(pred_rf,testing$classe)

cm$table

cm$overall["Accuracy"]
```
Accuracy is 99.6%, which means Out-of-sample error is only 0.4%, is pretty good

# Predict  #
Finally, we apply our model to offical test dataset. 
```{r Final Predict , echo=TRUE}
processed_test_data <- test_data[, names(test_data) %in% names(testing)]

final_predict <- predict(mod_rf, processed_test_data)
final_predict
```









